Funding programmes for early career researchers do/not influence their independence

### 2023-09-15 article.Rmd
edited tabels for readibility

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
library(tidyr)
library(dplyr)
library(cobalt)
library(knitr)
library(psych)
library(kableExtra)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(ez)
library(plm)

@@

When Looking at disciplines, we see that there are slightly more missing data in Agricultural and veterinary scieces and the Humanities and the Arts.

@@

# assuming df is your dataframe
df <- filtered_by_subclass %>% filter(independence_timing == "before_intervention")

# list of numeric columns. Exclude treatment column
numeric_vars <- df %>% select(where(is.numeric)) %>% select(-weights, -funded_control, -id, -treatment_year, -career_start_year) %>% names()

# descriptive stats for treatment level 1
table_before_1 <- df %>% filter(treatment == 1) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

# Print table for treatment level 1
kable(table_before_1, digits = 3, caption = "Descriptive statistics for treatment group") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# descriptive stats for treatment level 0
table_before_0 <- df %>% filter(treatment == 0) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

# Print table for treatment level 0
kable(table_before_0, digits = 3, caption = "Descriptive statistics for unfunded control group") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# descriptive stats for treatment level 2
table_before_2 <- df %>% filter(treatment == 2) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

# Print table for treatment level 2
kable(table_before_2, digits = 3, caption = "Descriptive statistics for funded control group") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

@@

# assuming df is your dataframe
df <- filtered_by_subclass %>% filter(independence_timing == "after_intervention")

# list of numeric columns. Exclude treatment column
numeric_vars <- df %>% select(where(is.numeric)) %>% select(-weights, -funded_control, -id, -treatment_year, -career_start_year) %>% names()

# descriptive stats for treatment level 1
table_before_1 <- df %>% filter(treatment == 1) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

# Print table for treatment level 1
kable(table_before_1, digits = 3, caption = "Descriptive statistics for treatment group") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# descriptive stats for treatment level 0
table_before_0 <- df %>% filter(treatment == 0) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

# Print table for treatment level 0
kable(table_before_0, digits = 3, caption = "Descriptive statistics for unfunded control group") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# descriptive stats for treatment level 2
table_before_2 <- df %>% filter(treatment == 2) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

# Print table for treatment level 2
kable(table_before_2, digits = 3, caption = "Descriptive statistics for funded control group") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

@@

BTW: there is an overlap between control group observations - `r overlap_control` observations from the unfunded control group also appears in the funded control group

@@

@@

summary(tar_read(matched_obj_unfunded))

plot(unf_sum, var.order = "unmatched")

cobalt::bal.tab(tar_read(matched_obj_unfunded), un = FALSE, stats = c("m", "v", "ks"), binary = "std")

cobalt::love.plot(tar_read(matched_obj_unfunded), binary = "std", var.order = "adjusted")

@@

summary(tar_read(matched_obj_funded))

plot(fun_sum, var.order = "unmatched")

@@

POZN: only for comparison:
To estimate the treatment effect and its standard error, we fit a linear regression model with 1978 earnings as the outcome and the treatment, covariates, and their interaction as predictors and included the full matching weights in the estimation. The lm() function was used to fit the outcome, and the comparisons() function in the marginaleffects package was used to perform g-computation in the matched sample to estimate the ATT. A cluster-robust variance was used to estimate its standard error with matching stratum membership as the clustering variable.

The estimated effect was XXXX (SE = XXXX, p = XXXX), indicating that the average effect of the treatment for those who received it is to increase earnings.

@@

filtered_by_subclass_1 <- pp %>% 
    group_by(subclass) %>% 
    filter(!any(treatment == 2)) %>% 
    ungroup()

@@

model <- lmer(RII ~ treatment * independence_timing + (1 | subclass), data = filtered_by_subclass_1)
summary(model)
r.squaredGLMM(model)

##checking assumptiosn fro repeated measures ANOVA

@@

model_lm <- lm(RII ~ treatment + subclass + subclass:independence_timing, data = filtered_by_subclass_1)

# Check residuals from lm
plot(resid(model_lm))

@@

plot(model_lm$fitted.values, resid(model_lm), 
     xlab = "Fitted values", ylab = "Residuals", 
     main = "Residuals vs. Fitted Values")
abline(h = 0, col = "red")

# Normality of Residuals: 
qqnorm(resid(model_lm))
qqline(resid(model_lm))

shapiro.test(resid(model_lm))

@@

library(car)
durbinWatsonTest(model_lm)

@@

vif(model_lm)

@@

plot(hatvalues(model_lm), rstudent(model_lm),
     xlab = "Leverage", ylab = "Studentized Residuals",
     main = "Residuals vs. Leverage")
abline(h = c(-2,2), col = "red", lty = 2)

@@

library(car)
leveneTest(RII ~ treatment, data = filtered_by_subclass_1)

@@

boxplot(RII ~ treatment, data = filtered_by_subclass_1)

@@

result <- ezANOVA(data = filtered_by_subclass_1,
                 dv = RII,
                 wid = subclass,
                 within = .(independence_timing),
                 between = treatment)
print(result)

@@

result_disc <- ezANOVA(data = filtered_by_subclass_1,
                 dv = RII,
                 wid = subclass,
                 within = .(independence_timing),
                 between = .(treatment, field))
print(result_disc)

@@

 ggplot(data = filtered_by_subclass_1, aes(x = treatment, y = RII, fill = independence_timing)) +
         geom_violin(scale = "width") +
         scale_fill_discrete(name = "Independence Timing", labels = c("Before intervention", "After intervention")) +
         xlab("Treatment Group") +
         ylab("RII Score") +
         ggtitle("Violin Plot of RII Scores by Treatment and Time") +
         theme_bw()

means_disc <- aggregate(RII ~ field, data = filtered_by_subclass_1, FUN = mean)

num_disc <- filtered_by_subclass_1 %>%
  group_by(field) %>%
  summarise(count = n(), .groups = "drop")

means_disc <- left_join(means_disc, num_disc, by = "field")

# Step 1: Reorder the levels of 'field'
means_disc$field <- with(means_disc, factor(field, levels=unique(field[order(RII)])))

# Step 2: Optionally, sort the dataframe based on the reordered levels
means_disc <- means_disc %>% arrange(field)

ggplot(data = means_disc, aes(x = reorder(c(1:6), RII), y = RII, fill = field)) +
  geom_bar(stat = "identity", position = "dodge") +

  # Add the number of observations on top of the bars
  geom_text(aes(label=count), position=position_dodge(width=0.9), vjust=-0.5) +

  labs(x = "Field", y = "Mean RII") +
  scale_fill_discrete(name = "Field") +
  theme_bw()

@@

In the first analysis, we examined the effects of treatment, independence_timing, and their interaction on the dependent variable, researcher independence (RII) for the treatment group and unfunded control group. There was a significant main effect of independence_timing on RII, F(`r result$ANOVA[2, 2]`, `r result$ANOVA[2, 3]`) = `r result$ANOVA[2, 4]`, p = `r result$ANOVA[2, 5]` indicating a significant change in the independence from before to after the intervention. The effect size (GES) for this effect was `r result$ANOVA[2, 7]`, suggesting a moderate effect.
Importantly, there was a significant interaction effect between treatment and independence_timing on independence, F(`r result$ANOVA[3, 2]`, `r result$ANOVA[3, 3]`) = `r result$ANOVA[3, 4]`, p = `r result$ANOVA[3, 5]`. This interaction suggests that the effect of the intervention on independence differed based on the treatment group. The effect size (GES) for this interaction was `r result$ANOVA[3, 7]`, indicating a small effect.

In the subsequent analysis, we further incorporated the factor of discipline (field) to explore if there were differences across disciplines in the observed effects. There was a significant main effect of field on independence, F(`r result_disc$ANOVA[2, 2]`, `r result_disc$ANOVA[2, 3]`) = `r result_disc$ANOVA[2, 4]`, p = `r result_disc$ANOVA[2, 5]`, indicating that there were significant differences in RII across the six discipline groups. On average, Agricultural and veterinary sciences showed lowest independence, while Social Sciences and Humanities showed almost twice as large independence, (though it is important to note that small sample might provide unstable estimates (----> i SHOULD REMOVE THEM FROM THE ANALYSIS SINCE THESE SEEM TO BE TOO FEW OBSERVATIONS --> MAYBE I SHOULD ALSO DO A SEPARATE ANOVA ONLY TAKING INTO ACCOUNT ALL DISC BEFORE INTEVENTION AND AFTER INTERVENTION SEPARATELY BUT FOR BOTH CONTROL GROUPS). The effect size (GES) for this was `r result_disc$ANOVA[2, 7]`, suggesting a moderate effect. The interaction between treatment and field was not statistically significant, F(`r result_disc$ANOVA[4, 2]`, `r result_disc$ANOVA[4, 3]`) = `r result_disc$ANOVA[4, 4]`, p = `r result_disc$ANOVA[4, 5]`, suggesting that the effect of treatment on independence was consistent across disciplines. The three-way interaction among treatment, independence_timing, and field was not statistically significant, F(`r result_disc$ANOVA[6, 2]`, `r result_disc$ANOVA[6, 3]`) = `r result_disc$ANOVA[6, 4]`, p = `r result_disc$ANOVA[6, 5]`. This implies that the interaction effect of treatment and independence_timing on independence remained consistent across different disciplines.

POZN ON MODEL CHOICE:

In Conclusion:

Both the mixed-effects model and the RM-ANOVA suggest a significant interaction between the treatment and the timing of the intervention. This means that the effect of the intervention differed between the treatment and control groups.

Both methods also indicate that the timing of the intervention (i.e., whether it's before or after) has a significant effect on the outcome RII.

The models are in agreement with their primary findings, which is a good sign. It suggests that the treatment has a distinct effect on RII after the intervention, different from that of the control group.

However, it's worth noting that while the methods reach similar conclusions, they approach the problem from different angles. The mixed-effects model incorporates the correlation within each subclass through its random effect, while the RM-ANOVA primarily focuses on the repeated measures nature of the data. Both models provide valuable insights, but the mixed-effects model might be a more robust choice for this particular dataset, given its explicit modeling of the subclass structure.

My take:

I understand the results of ANOVa, but dont understand the results of the mixed effects model very much. I also like that ANOVa actually provides p statistic and effect size statistic

@@

filtered_by_subclass_2 <- pp %>% 
    group_by(subclass) %>% 
    filter(!any(treatment == 0)) %>% 
    ungroup()

@@

model2 <- lmer(RII ~ treatment * independence_timing + (1 | subclass), data = filtered_by_subclass_2)
summary(model2)
r.squaredGLMM(model2)

result2 <- ezANOVA(data = filtered_by_subclass_2,
                 dv = RII,
                 wid = subclass,
                 within = .(independence_timing),
                 between = treatment)
print(result)

@@

result_disc2 <- ezANOVA(data = filtered_by_subclass_2,
                 dv = RII,
                 wid = subclass,
                 within = .(independence_timing),
                 between = .(treatment, field))
print(result_disc2)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
library(tidyr)
library(dplyr)
library(cobalt)
library(knitr)
library(psych)
library(kableExtra)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(ez)
library(plm)
library(effectsize)
library(nlme)

@@

When Looking at disciplines, we see that there are slightly more missing data in Agricultural and veterinary sciences and the Humanities and the Arts.

@@

df <- filtered_by_subclass %>% filter(independence_timing == "before_intervention")
numeric_vars <- df %>% 
  select(where(is.numeric)) %>% 
  select(-weights, -funded_control, -id, -treatment_year, -career_start_year) %>% 
  names()

# descriptive stats for each treatment level
table_before_1 <- df %>% filter(treatment == 1) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)
table_before_0 <- df %>% filter(treatment == 0) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)
table_before_2 <- df %>% filter(treatment == 2) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

colnames(table_before_0) <- paste0(colnames(table_before_0), "_0")
colnames(table_before_2) <- paste0(colnames(table_before_2), "_2")

combined_table_before <- cbind(table_before_1, table_before_0[, -1], table_before_2[, -1])

kable(combined_table_before, 
      digits = 3, align = "c", booktabs = TRUE) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  add_header_above(c(" " = 1, "Treatment Group" = 5, "Unfunded Control Group" = 4, "Funded Control Group" = 4))

@@

df <- filtered_by_subclass %>% filter(independence_timing == "after_intervention")
numeric_vars <- df %>% 
  select(where(is.numeric)) %>% 
  select(-weights, -funded_control, -id, -treatment_year, -career_start_year) %>% 
  names()

# descriptive stats for each treatment level
table_after_1 <- df %>% filter(treatment == 1) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)
table_after_0 <- df %>% filter(treatment == 0) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)
table_after_2 <- df %>% filter(treatment == 2) %>% select(all_of(numeric_vars)) %>% describe() %>% select(n, mean, sd, min, max) %>% round(., digits = 3)

colnames(table_after_0) <- paste0(colnames(table_after_0), "_0")
colnames(table_after_2) <- paste0(colnames(table_after_2), "_2")

combined_table_after <- cbind(table_after_1, table_after_0[, -1], table_after_2[, -1])

kable(combined_table_after, 
      digits = 3, align = "c", booktabs = TRUE) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  add_header_above(c(" " = 1, "Treatment Group" = 5, "Unfunded Control Group" = 4, "Funded Control Group" = 4))

@@

BTW: there is an overlap between control group observations: `r overlap_control` observations from the unfunded control group also appears in the funded control group

@@

##Unfunded group

@@

# summary(tar_read(matched_obj_unfunded))
#     
# plot(unf_sum, var.order = "unmatched")

aa <- cobalt::bal.tab(tar_read(matched_obj_unfunded), un = TRUE, stats = c("m", "v", "ks"), binary = "std")

# aa <- aa[!grepl("^disc_ford", names(aa))]

# aa <- aa[!grepl("^disc_ford", names(aa$Balance))]

bal_df <- aa$Balance

bal_df <- bal_df[!grepl("^disc_ford", rownames(bal_df)), ]

# Display the cleaned-up table with kable
kable(bal_df, "html", caption = "Balance Statistics For Unfunded Control Group") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

aa$Balance <- bal_df

cobalt::love.plot(aa, binary = "std", var.order = "adjusted")

##Funded group

@@

# summary(tar_read(matched_obj_funded))
#     
# plot(fun_sum, var.order = "unmatched")

bb <- cobalt::bal.tab(tar_read(matched_obj_funded), un = TRUE, stats = c("m", "v", "ks"), binary = "std")

# aa <- aa[!grepl("^disc_ford", names(aa))]

# aa <- aa[!grepl("^disc_ford", names(aa$Balance))]

bal_df_funded <- bb$Balance

bal_df_funded <- bal_df_funded[!grepl("^disc_ford", rownames(bal_df_funded)), ]

# Display the cleaned-up table with kable
kable(bal_df_funded, "html", caption = "Balance Statistics For Funded Control Group") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

bb$Balance <- bal_df_funded

cobalt::love.plot(bb, binary = "std", var.order = "adjusted")

@@

<!-- POZN: only for comparison: -->
<!-- To estimate the treatment effect and its standard error, we fit a linear regression model with 1978 earnings as the outcome and the treatment, covariates, and their interaction as predictors and included the full matching weights in the estimation. The lm() function was used to fit the outcome, and the comparisons() function in the marginaleffects package was used to perform g-computation in the matched sample to estimate the ATT. A cluster-robust variance was used to estimate its standard error with matching stratum membership as the clustering variable. -->

<!-- The estimated effect was XXXX (SE = XXXX, p = XXXX), indicating that the average effect of the treatment for those who received it is to increase earnings. -->

@@

filtered_by_subclass_1 <- pp %>% 
    group_by(subclass) %>% 
    filter(!any(treatment == 2)) %>% 
    ungroup() %>%
    mutate(treatment = factor(treatment, levels = c(0, 1)))

@@

model <- lmer(RII ~ treatment * independence_timing + (1 | subclass), data = filtered_by_subclass_1)
# summary(model)
# r.squaredGLMM(model)

model_no_interaction <- lmer(RII ~ treatment + independence_timing + (1 | subclass), data = filtered_by_subclass_1)

RSS_no_interaction <- sum(residuals(model_no_interaction)^2)
RSS_full <- sum(residuals(model)^2)
delta_RSS <- RSS_no_interaction - RSS_full
R2_interaction <- delta_RSS / RSS_no_interaction
# print(R2_interaction)

model_disc <- lmer(RII ~ treatment * independence_timing * field + 
              (1 + independence_timing|subclass), 
              data = filtered_by_subclass_1)

# summary(model_disc)

##checking assumptions fro repeated measures ANOVA

result <- aov(RII ~ treatment * independence_timing + Error(subclass/independence_timing), data=filtered_by_subclass_1)
# summary(result)
eta_sq_result <- eta_squared(result)

@@

# Check residuals from lm
# plot(resid(result))

@@

# plot(result$fitted.values, resid(result), 
#      xlab = "Fitted values", ylab = "Residuals", 
#      main = "Residuals vs. Fitted Values")
# abline(h = 0, col = "red")

# Normality of Residuals: 
# qqnorm(resid(result))
# qqline(resid(result))
# 
# shapiro.test(resid(result))

@@

# library(car)
# durbinWatsonTest(result)

@@

# vif(result)

@@

# plot(hatvalues(result), rstudent(result),
#      xlab = "Leverage", ylab = "Studentized Residuals",
#      main = "Residuals vs. Leverage")
# abline(h = c(-2,2), col = "red", lty = 2)

@@

# library(car)
# leveneTest(RII ~ treatment, data = filtered_by_subclass_1)

@@

# boxplot(RII ~ treatment, data = filtered_by_subclass_1)

@@

# filtered_by_subclass_1 <- filtered_by_subclass_1 %>%
#      group_by(vedidk, subclass) %>%  # Assuming 'id' is your current unique identifier
#      mutate(subject_id = cur_group_id()) %>%  # This creates a unique ID for each group
#      ungroup()

# #this is preferable because it takes into account the matched nature of the data, as well as repeated measures nature of the data -> see chatGPT conversation "ANOVA `wid` parameter clarification"
# result <- aov(RII ~ treatment * independence_timing + Error(subclass/independence_timing), data=filtered_by_subclass_1)
# # summary(result)
# eta_sq_result <- eta_squared(result)

#this is less preferable because it takes into account only the repeated measure nature of the data, not the matched nature -> you can see that when you run this tests, there is warning that says "there were non-unique subclass values across the between-subjects variable (treatment). Essentially, ezANOVA detected the problem and internally "fixed" it by making subclass unique across the treatment groups, so it behaves similarly to subject_id you later created." -> this means that it effectively converts the subclass to subject_id anyway
# result <- ezANOVA(data = filtered_by_subclass_1,
#                  dv = RII,
#                  wid = subclass,
#                  within = .(independence_timing),
#                  between = treatment)
# # print(result)

@@

filtered_data_disc_1 <- filtered_by_subclass_1 %>%
  group_by(field) %>%
  filter(n() >= 10)

filtered_data_disc_1 <- filtered_data_disc_1 %>%
     group_by(vedidk, subclass) %>%  # Assuming 'id' is your current unique identifier
     mutate(subject_id = cur_group_id()) %>%  # This creates a unique ID for each group
     ungroup()

# Calculate weights based on group size (you can adjust this based on other criteria if needed)
group_sizes <- table(filtered_data_disc_1$field)
max_size <- max(group_sizes)
filtered_data_disc_1$weights <- max_size / group_sizes[filtered_data_disc_1$field]

#this model addresses both unequal group sizes (via weights) and heteroscedasticity (via varIdent). Additionally, I've added a first-order autoregressive structure (corAR1) assuming you have repeated measures on the same subjects and to account for their paired nature (indicated by subclass/independence_timing)
#this model directly tests whether there is a difference in disciplines - anyway it tests each discipline agains each other, so it doesnt produce the neat result of "yes, there are some significant differences between disciplines", but more nuanced version of "there is specifically difference between this and this discipline"

model_gls_weighted <- gls(RII ~ field, data = filtered_data_disc_1, 
                          weights = varIdent(form = ~ 1 | field),
                          method = "ML",
                          correlation = corAR1(form = ~ 1 | subject_id))

# summary(model_gls_weighted)

#this model below is the same as the above but with added predictors

model_gls_interactions_subclass <- gls(RII ~ field * treatment * independence_timing, 
                              data = filtered_data_disc_1, 
                              weights = varIdent(form = ~ 1 | field),
                              method = "ML",
                              correlation = corAR1(form = ~ 1 | subclass/independence_timing))

# summary(model_gls_interactions_subclass)

#one potential strategy to get the neat result of "yes, there are some significant differences between disciplines" is following: create two similar models

model_without_field <- gls(RII ~ treatment + independence_timing + treatment:independence_timing, 
                           data = filtered_data_disc_1, 
                           correlation = corCompSymm(form = ~ 1 | subclass/independence_timing), 
                           method = "ML")

model_with_field_main_effect <- gls(RII ~ treatment + field + independence_timing + treatment:independence_timing, 
                                    data = filtered_data_disc_1, 
                                    correlation = corCompSymm(form = ~ 1 | subclass/independence_timing), 
                                    method = "ML")

anova_result <- anova(model_without_field, model_with_field_main_effect)

# Your result specifically shows that the model_with_field_main_effect is significantly better at explaining the variance in your response variable compared to the model_without_field. The p-value is extremely small (1e-04), indicating a significant main effect of "field".
# 
# In other words, the main effect of "field" significantly improves the model fit, suggesting that the variable "field" has a significant main effect on the response variable RII after accounting for other predictors in the model.

#this model does NOT address unequal group sizes (via weights) and heteroscedasticity

# result_disc <- ezANOVA(data = filtered_data_disc_1,
#                  dv = RII,
#                  wid = subclass,
#                  within = .(independence_timing),
#                  between = .(treatment, field))
# # print(result_disc)

In the first analysis, we examined the effects of treatment, independence_timing, and their interaction on the dependent variable, researcher independence (RII) for the treatment group and unfunded control group.

There was a significant main effect of independence_timing on RII, F(`r summary(result)$"Error: subclass:independence_timing"[[1]]$"Df"[1]`, `r format(summary(result)$"Error: subclass:independence_timing"[[1]]$"Df"[2])`) = `r format(summary(result)$"Error: subclass:independence_timing"[[1]]$"F value"[1], digits=2)`, p = `r format(summary(result)$"Error: subclass:independence_timing"[[1]]$"Pr(>F)"[1], digits=4)`). This effect accounted for approximately `r format(eta_sq_result[eta_sq_result$Parameter == "independence_timing", "Eta2_partial"] * 100, digits=2)`% of the observed variance (partial \(\eta^2 = `r eta_sq_result[eta_sq_result$Parameter == "independence_timing", "Eta2_partial"]`\), 95% CI [`r eta_sq_result[eta_sq_result$Parameter == "independence_timing", "CI_low"]`, `r eta_sq_result[eta_sq_result$Parameter == "independence_timing", "CI_high"]`]).

Importantly, there was a significant interaction effect between treatment and independence_timing on independence, F(`r format(summary(result)$"Error: Within"[[1]]$"Df"[2])`, `r format(summary(result)$"Error: Within"[[1]]$"Df"[3])` = `r format(summary(result)$"Error: Within"[[1]]$"F value"[2], digits=2)`, p = `r format(summary(result)$"Error: Within"[[1]]$"Pr(>F)"[2], digits=2)`. This interaction effect explained about `r format(eta_sq_result[eta_sq_result$Parameter == "treatment:independence_timing", "Eta2_partial"] * 100, digits=2)`% of the variability (partial \(\eta^2 = `r eta_sq_result[eta_sq_result$Parameter == "treatment:independence_timing", "Eta2_partial"]`\), 95% CI [`r eta_sq_result[eta_sq_result$Parameter == "treatment:independence_timing", "CI_low"]`, `r eta_sq_result[eta_sq_result$Parameter == "treatment:independence_timing", "CI_high"]`]).

<!-- LMM interpretation: -->

<!-- In the updated analysis using a linear mixed model, we examined the effects of treatment, independence_timing, and their interaction on the dependent variable, researcher independence (RII) for the treatment group and unfunded control group. -->

<!-- - There was a significant main effect of `independence_timing` on RII, t(`r summary(model)$coefficients[3,3]`) = `r summary(model)$coefficients[3, "t value"]`, p = `r format(summary(model)$coefficients[3,5], scientific = TRUE)`, suggesting that there was a significant change in RII from before to after intervention in the control group. -->

<!-- - There was also a significant main effect of `treatment` on RII, t(`r summary(model)$coefficients[2,3]`) = `r summary(model)$coefficients[2, "t value"]`, p = `r format(summary(model)$coefficients[2,5], scientific = TRUE)`. This effect indicates a difference in RII between the control group and the treatment group before the intervention. -->

<!-- - Importantly, there was a significant interaction effect between `treatment` and `independence_timing` on RII, t(`r summary(model)$coefficients[4,3]`) = `r summary(model)$coefficients[4, "t value"]`, p = `r format(summary(model)$coefficients[4,5], scientific = TRUE)`. This interaction suggests that the effect of the intervention on independence differed based on the timing of the measurement. The effect size for all fixed effects can be approximated using the marginal \( R^2 \) which is `r r.squaredGLMM(model)[1]`, suggesting that about `r round(r.squaredGLMM(model)[1] * 100, 1)`% of the variance in RII can be explained by the main effects in our model. The effect size for the interaction specifically can be computed using \( \Delta R^2 \) due to the interaction, which is `r R2_interaction`, indicating that about `r round(R2_interaction * 100, 1)`% of variance in RII is explained by the interaction between treatment and independence_timing over and above the main effects. -->

<!-- POZN ON MODEL CHOICE: -->

<!-- In Conclusion: -->

<!-- Both the mixed-effects model and the RM-ANOVA suggest a significant interaction between the treatment and the timing of the intervention. This means that the effect of the intervention differed between the treatment and control groups. -->

<!-- Both methods also indicate that the timing of the intervention (i.e., whether it's before or after) has a significant effect on the outcome RII. -->

<!-- The models are in agreement with their primary findings, which is a good sign. It suggests that the treatment has a distinct effect on RII after the intervention, different from that of the control group. -->

<!-- However, it's worth noting that while the methods reach similar conclusions, they approach the problem from different angles. The mixed-effects model incorporates the correlation within each subclass through its random effect, while the RM-ANOVA primarily focuses on the repeated measures nature of the data. Both models provide valuable insights, but the mixed-effects model might be a more robust choice for this particular dataset, given its explicit modeling of the subclass structure. -->

<!-- My take: -->

<!-- The advantage of ANOVa is that it provides information about the "average" effect of independence timing across the dataset, whereas LMM does not do that (but it could be calculated extra, by default it only shows change in time in the control group, not the treatment group, which makes the interpretation harder). On the other hand  -->
<!-- None of these models work well with the disciplines by default - I could make some extra edits to make ANOVA work, but I'm unsure what the interpretation would be for LMM -->

@@

 # ggplot(data = filtered_by_subclass_1, aes(x = treatment, y = RII, fill = independence_timing)) +
 #         geom_violin(scale = "width") +
 #         scale_fill_discrete(name = "Independence Timing", labels = c("Before intervention", "After intervention")) +
 #         xlab("Treatment Group") +
 #         ylab("RII Score") +
 #         ggtitle("Violin Plot of RII Scores by Treatment and Time") +
 #         theme_bw()

 # Calculate means and standard errors for plotting
plot_data <- filtered_by_subclass_1 %>%
  group_by(treatment, independence_timing) %>%
  summarize(
    mean_RII = mean(RII),
    se_RII = sd(RII) / sqrt(n())
  )

# # Plot
# p <- ggplot(plot_data, aes(x = independence_timing, y = mean_RII, group = treatment, color = factor(treatment))) +
#   geom_line(aes(linetype = factor(treatment))) + 
#   geom_point(size = 3) + 
#   geom_errorbar(aes(ymin = mean_RII - se_RII, ymax = mean_RII + se_RII), width = 0.2) +
#   labs(
#     title = "Effect of Intervention on Researcher Independence Index (RII)",
#     x = "Independence Timing",
#     y = "Mean RII",
#     color = "Treatment Group",
#     linetype = "Treatment Group"
#   ) +
#   scale_color_manual(values = c("blue", "red"), labels = c("Control", "Treatment")) +
#   scale_linetype_manual(values = c("dashed", "solid")) +
#   theme_minimal()

dodge_width <- 0.5

combined_plot <- ggplot() +

  geom_violin(data = filtered_by_subclass_1, 
              aes(x = treatment, y = RII, fill = independence_timing), 
              scale = "width", 
              position = position_dodge(dodge_width)) +

  geom_point(data = plot_data, 
             aes(x = treatment, y = mean_RII, color = independence_timing), 
             position = position_dodge(dodge_width), 
             size = 3) +

  geom_errorbar(data = plot_data, 
                aes(x = treatment, ymin = mean_RII - se_RII, ymax = mean_RII + se_RII, color = independence_timing), 
                position = position_dodge(dodge_width), 
                width = 0.2) +

  scale_fill_discrete(name = "Independence Timing", labels = c("Before intervention", "After intervention")) +
  scale_color_manual(values = c("black", "red"), name = "Independence Timing", labels = c("Before intervention", "After intervention")) +

  labs(title = "Violin Plot of RII Scores by Treatment and Time with Mean and CI Overlayed",
       x = "Treatment Group",
       y = "RII Score") +
  theme_bw()

print(combined_plot)

@@

@@

# means_disc <- aggregate(RII ~ field, data = filtered_by_subclass_1, FUN = mean)
#      
# num_disc <- filtered_by_subclass_1 %>%
#   group_by(field) %>%
#   summarise(count = n(), .groups = "drop")
# 
# means_disc <- left_join(means_disc, num_disc, by = "field") 
# 
# # Step 1: Reorder the levels of 'field'
# means_disc$field <- with(means_disc, factor(field, levels=unique(field[order(RII)])))
# 
# # Step 2: Optionally, sort the dataframe based on the reordered levels
# means_disc <- means_disc %>% arrange(field)
# 
#     
# se_ci_data <- filtered_by_subclass_1 %>%
#   group_by(field) %>%
#   summarise(mean_RII = mean(RII),
#             se_RII = sd(RII) / sqrt(n()),
#             lower_CI = mean_RII - (1.96 * se_RII),
#             upper_CI = mean_RII + (1.96 * se_RII), 
#             .groups = "drop")
# 
# means_disc <- merge(means_disc, se_ci_data, by = "field")
# 
# 
# ggplot(data = means_disc, aes(x = field, y = mean_RII, fill = field)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   
#   # Add error bars for confidence intervals
#   geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = position_dodge(width=0.9), width=0.25) +
#   
#   # Add the number of observations in the middle of the bars
#   geom_text(aes(label=count, y = mean_RII/2), position=position_dodge(width=0.9), vjust=0.5) +
#   
#   # Change x-axis labels and add a title
#   scale_x_discrete(labels = seq_along(unique(means_disc$field))) + # replace with numbers
#   labs(
#     x = "Field", 
#     y = "Mean RII",
#     title = "Differences in RII across disciplines" # Replace with your desired title
#   ) +
#   scale_fill_discrete(name = "Field") +
#   theme_bw()

filtered_by_subclass_2 <- pp %>% 
    group_by(subclass) %>% 
    filter(!any(treatment == 0)) %>% 
    ungroup() %>%
    mutate(treatment = as.numeric(as.character(treatment))) %>%
    mutate(treatment = if_else(treatment == 2, 0, treatment)) %>%
    mutate(treatment = factor(treatment, levels = c(0, 1)))

@@

# model2 <- lmer(RII ~ treatment * independence_timing + (1 | subclass), data = filtered_by_subclass_2)
# summary(model2)
# r.squaredGLMM(model2)

result2 <- aov(RII ~ treatment * independence_timing + Error(subclass/independence_timing), data=filtered_by_subclass_2)
# summary(result)
eta_sq_result2 <- eta_squared(result2)

# result2 <- ezANOVA(data = filtered_by_subclass_2,
#                  dv = RII,
#                  wid = subclass,
#                  within = .(independence_timing),
#                  between = treatment)
# print(result)

@@

# result_disc2 <- ezANOVA(data = filtered_by_subclass_2,
#                  dv = RII,
#                  wid = subclass,
#                  within = .(independence_timing),
#                  between = .(treatment, field))
# print(result_disc2)

In the second analysis, we examined the effects of treatment, independence_timing, and their interaction on the researcher independence (RII) for the treatment group and funded control group. Again, there was a significant main effect of time on independence, F(`r format(summary(result2)$"Error: subclass:independence_timing"[[1]]$"Df"[1])`, `r format(summary(result2)$"Error: subclass:independence_timing"[[1]]$"Df"[2])`) = `r format(summary(result2)$"Error: subclass:independence_timing"[[1]]$"F value"[1], digits=2)`, p = `r format(summary(result2)$"Error: subclass:independence_timing"[[1]]$"Pr(>F)"[1], digits=4)`). The effect accounted for approximately `r format(eta_sq_result2[eta_sq_result2$Parameter == "independence_timing", "Eta2_partial"] * 100, digits=2)`% of the observed variance (partial \(\eta^2 = `r eta_sq_result2[eta_sq_result2$Parameter == "independence_timing", "Eta2_partial"]`\), 95% CI [`r eta_sq_result2[eta_sq_result2$Parameter == "independence_timing", "CI_low"]`, `r eta_sq_result2[eta_sq_result2$Parameter == "independence_timing", "CI_high"]`]).
